### 2024-08-23 sandbox.Rmd
file upload

---
title: "data sandbox"
output: html_document
date: "2024-07-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Tratando de ver que datos tengo

```{r world development indicators}
colnames(wdi) <- c("C_Name","C_Code","S_Name","S_Code","YR1995","YR1996","YR1997","YR1998", "YR1999","YR2000","YR2001","YR2002","YR2003","YR2004","YR2005","YR2006","YR2007","YR2008","YR2009","YR2010","YR2011","YR2012","YR2013","YR2014","YR2015","YR2016","YR2017","YR2018","YR2019","YR2020","YR2021","YR2022","YR2023")
```

```{r}
wdi %>% 
  filter(C_Code == "AFG") %>%
  select()
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.